Cooperative Kernels: Safe Multitasking for Blocking Algorithms on GPUs

### 2016-11-11 write-ups/PLDI/paper.tex
More semantic differences.

Semantically, when all threads have reached $\resizingglobalbarrier$,
the number of active workgroups, $M$, is set to a new value, $M'$ say.
If $M' = M$ then the active workgroups remain unchanged and proceed to
execute after the barrier.  If $M' < M$, workgroups $[M', M-1]$ are
killed.  If $M' > M$ then $M'-M$ new workgroups join the computation,
as if they were forked from workgroup 0.  In particular, the
$\transmit$-annotated local state of thread 0 in workgroup 0 is
transmitted to the threads of the new workgroups.  The change in
number of active workgroups from $M$ to $M'$ is atomic.

@@

\ADComment{TODO: To simplify programming, avoiding the need to wrap all kills in conditionals, we always keep one workgroup alive.  This fits well with our decision to kill WGs in decreasing order.}

\ADComment{TODO: There are possible use cases for allowing all WGs to be killed, though, and we might wish to allow them.}

@@

\ADComment{Something old to maybe work in:}

The programmer is responsible for keeping at least one workgroup alive.  In particular, if all workgroups execute $\offerkill$, the scheduler is at liberty to destroy all workgroups, causing the cooperative kernel to exit.  This can be guarded against by wrapping each $\offerkill$ call in a conditional statement:

\lstset{basicstyle=\tt,numbers=none}
\begin{lstlisting}
  if($\getgroupid{0}$ != 0) $\offerkill$();
\end{lstlisting}
\lstset{basicstyle=\scriptsize\tt,numbers=left}

We could, alternatively, have formulated the semantics of $\offerkill$
so that workgroup 0 is never killed.  However, there might be a use
case for a cooperative kernel reaching a point where it would be
acceptable for the kernel to exit, although desirable for some
remaining computation to be performed if competing workloads allow it.

>>>>

Semantically, when all threads have reached $\resizingglobalbarrier$,
the number of active workgroups, $M$, is set to a new value, $M'$ say, with $M' > 0$.
If $M' = M$ then the active workgroups remain unchanged and proceed to
execute after the barrier.  If $M' < M$, workgroups $[M', M-1]$ are
killed.  If $M' > M$ then $M'-M$ new workgroups join the computation,
as if they were forked from workgroup 0.  In particular, the
$\transmit$-annotated local state of thread 0 in workgroup 0 is
transmitted to the threads of the new workgroups.  The change in
number of active workgroups from $M$ to $M'$ is atomic.

@@

Our semantics dictate that the workgroup with id 0 will not be killed
if it invokes $\offerkill$.  This avoids the possibility of the
cooperative kernel terminating early due to the programmer
inadvertantly allowing all workgroups to be killed, and the decision
to keep workgroup 0 alive fits well with our choice to kill workgroups
in descending order of id.

However, there might be a use case for a cooperative kernel reaching a
point where it would be acceptable for the kernel to exit, although
desirable for some remaining computation to be performed if competing
workloads allow it.  In this case, a semantics where all workgroups can be killed via $\offerkill$ would be appropriate, and the programmer would need to guard each $\offerkill$ with an id check in cases where killing all workgroups would be unacceptable.  For example:
%
\lstset{basicstyle=\tt,numbers=none}
\begin{lstlisting}
  if($\getgroupid{0}$ != 0) $\offerkill$();
\end{lstlisting}
\lstset{basicstyle=\scriptsize\tt,numbers=left}
%
would ensure that at least workgroup 0 is kept alive.

@@
